Querying Integrated Scientific Observation and Measurement Data

### 2010-04-30 docs/algdraft/DataIntegration/DataIntegration.tex
add more on the materialized db part.

\begin{itemize}
\item $Q_1$: Give me the datasets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the datasets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item $Q_4$: Give me the datasets that have trees with average ``height''
  than 20in in year 2009. 
\item $Q_5$: an example on summarize one measure with respect to another.
This could take two forms: one where the context forms the groups
(like by year), and another where the two measurements are within the
same context (like finding correlations).
\end{itemize}

@@

\section{Naive method}\label{sec:naive}
Index the data with all the values, then search over values. 
This may get wrong results.

@@

\subsection{Framework and operators}

@@

The previous section shows the method to rewrite a given query. 
In this section, we would discribe another strategy.

In this method, we merge the data in different tables into several
core OBOE tables, denoted by $OBOE.*$
In detail,
\begin{itemize}
\item $OBOE.EntityIns=\{\langle \underline{EntId}, EntType\rangle\}$ for all the entity instances;
\item $OBOE.ObsIns=\{\langle \underline{ObsId}, ObsType, EntId \rangle\}$ keeps all the observation instances;
\item $OBOE.MeasIns=\{\langle \underline{MeasId}, ObsId, MeasType, Val\rangle\}$ for all the measurement instances;
\item $OBOE.ContextIns =\{\langle \underline{ObsId, ConextObsId, ContextType}\rangle\}$ for all the context instances;
\end{itemize}

\subsubsection{Materialize data}
{\bf: TODO: refine the writing in MaterializeDB and move them here.}

\subsubsection{Querying materialized database}

>>>>

\begin{itemize}
\item $Q_1$: Give me the datasets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the datasets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item $Q_4$: Give me the datasets that have trees with average ``height''
  than 20in in year 2009. 
% Shawn: an example on summarize one measure with respect to another.
% This could take two forms: one where the context forms the groups
% (like by year), and another where the two measurements are within the
% same context (like finding correlations).
\end{itemize}

@@

\section{Naive method}\label{sec:naive}
A very direct way to help searching would be to index the data with
the data with all the values, then search over these values. 
This may get wrong results.

@@

%\subsection{Framework and operators}

@@

The previous section shows the method to rewrite a given query. 
However, as we analyzed, the computation cost may be higher when the
candidate data tables are many. In this section, we propose another
strategy to make use of the existing optimization strategies of
current DBMS.

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In partibular, we have the following concept instance tables. 
\begin{itemize}
\item $OBOE.EntityIns=\{\langle \underline{EntId}, EntType\rangle\}$ for all the entity instances;
\item $OBOE.ObsIns=\{\langle \underline{ObsId}, ObsType, EntId \rangle\}$ keeps all the observation instances;
\item $OBOE.MeasIns=\{\langle \underline{MeasId}, ObsId, MeasType, Val\rangle\}$ for all the measurement instances;
\item $OBOE.ContextIns =\{\langle \underline{ObsId, ConextObsId, ContextType}\rangle\}$ for all the context instances;
\end{itemize}

In what follows, we describe how to materialize the data into such
concept instance tables in Section \ref{sec:materialilzedb} when
considering all the distinct and context constratins. 
Then, in Section \ref{sec:materialilzedb}, we describe our strategy to
perform queries over materialized database.

\subsubsection{Materialize data}\label{sec:materialilzedb}
{\bf: TODO: refine the writing in MaterializeDB and move them here.}

\subsubsection{Querying materialized database}\label{sec:materialilzedb}
